Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-02-06 thesis.qmd
quick_comp+units+N_sample
SFR-comparison

### Estimating SFR from Spectra

@@

-   **Hα** \boldmath$\alpha$ **Emission**: Young, hot, massive stars (O-type stars, \~10 Myr, \~20 $M_\odot$) produce a number of ionizing photons, which they ionize the surrounding hydrogen rich gas. The hydorgen undergoes recombination cascades which produce Balmer emission lines of $H\alpha$ (0.6563 $\mu m$) and $H\beta$ (0.4861 $\mu m$). Dust can significantly affect observations.

@@

![Plot showing the linear relation $\log_{10}\text{SFR}_{FUV} = \log_{10}\text{SFR}_{H\alpha}$, as well as their distrubutions](figure/log_SFR_FUV_Ha.png){fig-align="center" width="500"}

@@

$$
SFR_{\text{total}} = \text{mean}\left(\mathcal{K}_{\text{FUV}} \cdot L_{\text{FUV}}, \mathcal{K}_{\text{H}\alpha} \cdot L_{\text{H}\alpha}\right) + \mathcal{K}_{\text{IR}} \cdot L_{\text{IR}} \\
= SFR_{\text{FUV},H\alpha}+ SFR_{\text{IR}}
$$

@@

-   Distances \[Mpc\] (288/288 galaxies)
-   Radial Velocities \[km/s\] (286/288 galaxies)
-   Morphology and Geometry of the galaxies
    -   Morphological type (229/288 galaxies)
    -   Inclination \[degrees\] (209/288 galaxies)
    -   Angular Diameter of the Major and Semi-major axis \[arcmin\] (261/288 galaxies)
-   K-band Luminosity \[$L_\odot$\] and Magnitude (70/288 galaxies)
-   B-band Magnitudes \[mag\] (244/288 galaxies)
-   Star Formation Rates \[$M_\odot/yr$\] (73/288 galaxies)

For the comparison of our data we will use linear regression, accounting for the errors of the observed data, where $R^2$ and the slope of the fitted line should be close to 1[^6]. We will also plot the residuals ($y_\text{Regression}-y_\text{observed}$) and if the mean value $\overline{\text{Residuals}}\approx 0$, then the linear regression is succesfull.

[^6]: $R^2$ Measures the proportion of variance explained by the linear model. For a model to be acceptable the $R^2$ should be higher than $80\%$, or close to it if $\overline{\text{Residuals}}\approx 0$

All the results are shown in @sec-regressions, except the comparisons of the Morphological Types and of the SFRs,who need more detailed analysis.

To begin with the morphological types of the galaxies, we first need to address possible potential inconsistencies in the classifications of the two catalogs. As we can see in @fig-TType-all, some galaxies "jump" from one side of the classification to the other side, depending on the catalog. These galaxies are dwarf galaxies, so we can safely assume that they are Tr type galaxies, as discussed in @sec-dwarves. For the rest of the analysis we will adopt the Numerical Morphological Type from the HECATE catalog, since it includes a larger number of galaxies, ensuring better
data compatibility. Additionally, the HECATE catalog boasts an accuracy of 0.1 and provides errors in comparison with the UNGC.

@@

>>>>

### Estimating SFR from Spectra {#sec-SFR-est}

@@

-   **H**\boldmath$\alpha$ **Emission**: Young, hot, massive stars (O-type stars, \~10 Myr, \~20 $M_\odot$) produce a number of ionizing photons, which they ionize the surrounding hydrogen rich gas. The hydorgen undergoes recombination cascades which produce Balmer emission lines of $H\alpha$ (0.6563 $\mu m$) and $H\beta$ (0.4861 $\mu m$). Dust can significantly affect observations.

@@

![Plot showing the linear relation $\log_{10}\text{SFR}_{FUV} = \log_{10}\text{SFR}_{H\alpha}$, as well as their distributions, based on the data from UNGC](figure/log_SFR_FUV_Ha.png){#fig-SFR-Ha-FUV fig-align="center" width="500"}

@@

$$
SFR_{\text{total}} = \text{mean}\left(\mathcal{K}_{\text{FUV}} \cdot L_{\text{FUV}}, \mathcal{K}_{\text{H}\alpha} \cdot L_{\text{H}\alpha}\right) + \mathcal{K}_{\text{IR}} \cdot L_{\text{IR}} \\
= SFR_{\text{FUV},H\alpha}+ SFR_{\text{IR}}
$$ {#eq-SFR-tot-mean}

@@

-   Distances \[Mpc\] (288/288 common galaxies)
-   Radial Velocities \[km/s\] (286/288 common galaxies)
-   Morphology and Geometry of the galaxies
    -   Morphological type (229/288 common galaxies)
    -   Inclination \[degrees\] (209/288 common galaxies)
    -   Angular Diameter of the Major and Semi-major axis \[arcmin\] (261/288 common galaxies)
-   K-band Luminosity \[$L_\odot$\] and Magnitude (70/288 common galaxies)
-   B-band Magnitudes \[mag\] (244/288 common galaxies)
-   Star Formation Rates \[$M_\odot/yr$\] (73/288 common galaxies)

For the comparison of our data we will use linear regression, accounting for the errors of the observed data, where $R^2$ and the slope of the fitted line should be close to 1[^6]. We will also plot the residuals ($y_\text{Regression}-y_\text{observed}$) and if the mean value $\overline{\text{Residuals}}\approx 0$, then the linear regression is considered succesfull.

[^6]: $R^2$ measures the proportion of variance explained by the linear model. For a model to be acceptable the $R^2$ should be higher than $80\%$, or close to it if $\overline{\text{Residuals}}\approx 0$

All the results are shown in @sec-regressions, except the comparisons of the Morphological Types and of the SFRs, that need more detailed analysis.[^7]

[^7]: The names of the x and y axes are written as they are given in the corresponding catalog, except the ones in decimal logarithmic scale, that are notated appropriately.

To begin with the morphological types of the galaxies, we first need to address potential inconsistencies in the classifications of the two catalogs. As we can see in @fig-TType-all, some galaxies "jump" from one side of the classification to the other side, depending on the catalog. These galaxies are dwarf galaxies, so we can safely assume that they are Tr type galaxies, as discussed in @sec-dwarves. For the rest of the analysis we will adopt the Numerical Morphological Type from the HECATE catalog, since it includes a larger number of galaxies, ensuring better data compatibility. Additionally, the HECATE catalog boasts an accuracy of 0.1 and provides errors, while the UNGC provides a rigid categorization of the galaxies with no errors.

@@

Next we need to analyze the SFR from the two catalogs. The catalogs provide different SFR estimators, H$\alpha$ and FUV from UNGC and the total IR from HECATE. As already discussed in @sec-SFR-est, the two estimators FUV and H$\alpha$ can be combined using their average ($\text{mean}(SFR_{FUV},SFR_{H\alpha})$), however both estimators and their average can potentially underestimate the SFR due to dust attenuation ($SFR_{TIR}$ is not affected by this), and this is exactly what is observed in @fig-SFR-comparison.

![Scatter plot and the residuals of the $\text{mean}(SFR_{FUV},SFR_{H\alpha})$ from the UNGC catalog and the $SFR_{TIR}$. The SFR estimation from the FUV and H$\alpha$ is systematically lower than the $SFR_{TIR}$ estimator as expected. The $R^2$ is smaller than 80% however due to $\overline{\text{Residuals}}=0$ we can assume that the quantities are comparable.](images/paste-1.png){#fig-SFR-comparison width="60%"}

In reality those estimators compliment each other and this is exactly why we they are not equal. The total SFR could potentially be calculated by @eq-SFR-tot-mean, however due to the lack of samples with both traces the calculation of $\text{SFR}_\text{total}$ is not possible without introducing biases to our data. Due to the underestimation from the estimators of UNGC, the main estimator used in our analysis will be the $TIR$ estimator and the $\text{mean}(SFR_{FUV},SFR_{H\alpha})$ will be transformed into the expected $\text{SFR}_\text{TIR}$ according to the linear regression of @fig-SFR-comparison.
